Outline paper - Multiple imputation for statistical disclosure control: Creating synthetic datasets with *mice*

### 2021-09-25 mice_synthesizing/paper.Rmd
Minor revisions.

Open science, including open data, has been marked as the future of science [@gewin_data_2016], and the advantages of publicly available research data are numerous [@molloy_open_2011; @walport_brest_sharing_2011]. Collecting research data requires an enormous investment both in terms of time and monetary resources. Openly accessible research data bears the potential of increasing the scientific returns for the same data collection effort. Additionally, the fact that public funds are used for data collection results in increasing demand for the collected data. Nevertheless, the possibilities to distribute research data directly are often very limited due to restrictions on data privacy and data confidentiality. Although these regulations are much needed, privacy constraints are also ranked among the toughest challenges to overcome in the advancement of computational social science [@lazer_life_2009].

Anonymizing the data might seem a quick and appealing approach to limit the unique identification of participants. However, this approach is not sufficient to fulfill contemporary privacy and confidentiality requirements [@ohm_broken_2009; @national_putting_2007]. Over the years, several other techniques have been used to increase the confidentiality of the data, such as categorizing continuous variables, top coding values above an upper bound or adding random noise to the observed values [@drechsler_synthetic_2011]. However, these methods may distort the true data relation between variables, thereby reducing the data quality and the scientific returns for re-using the same data for further research.

@@

Conceptually, the synthetic data framework is based upon the building blocks of multiple imputation of missing data, as proposed by @rubin_multiple_1987. Instead of replacing just the missing values with multiple draws from the posterior predictive distribution, one could easily *overimpute* any observed sensitive values. Similarly to multiple imputation of missing data, the multiple synthetic datasets allow for correct statistical inferences, **regardless of the fact that the "true" value is not included in the analyses**. The analyses over multiple synthetic data sets should be pooled into a single inference, so that the researcher can draw valid conclusions from the pooled results. To that respect, the variance should reflect the added variability that is induced by the imputation procedure. 
Potentially, this approach could fulfill the needs for openly accessibly data, without running into barriers with regard to privacy and confidentiality constraints. However, there is no such thing as a free lunch: data collectors have to put effort in creating high-quality synthetic data. Also, the quality of the synthetic data is highly dependent on the imputation models, and using flawed models to generate synthetic data might bias subsequent analyses. Conversely, if the models used to create the synthetic data are able to preserve the relationships between the variables as in the original data, the synthetic data can be nearly as informative as the observed data. Thus, to fully exploit the benefits of synthetic data, the effort to actually create these high-quality data sets should be kept at a minimum.

@@

For example, overimputing synthetic values for both the observed and missing cells in the `mice::boys` data set into the object `syn`, given the same imputation parameters as before, can be realized by the following code execution

@@

# Simulation setup
- boys data met bootstrap om sampling variance er in te brengen
- wat evalueren we: bias | standard error | coverage
- welke pooling: T = Ubar + B/m

# psych::describe - univariate ok
map --> psych::describe() --> Reduce("+", .) in de pipe

>>>>

Open science, including open data, has been marked as the future of science [@gewin_data_2016], and the advantages of publicly available research data are numerous [@molloy_open_2011; @walport_brest_sharing_2011]. Collecting research data requires an enormous investment both in terms of time and monetary resources. Openly accessible research data bears the potential of increasing the scientific returns for the same data collection effort. Additionally, the fact that public funds are used for data collection results in increasing demand for the collected data. Nevertheless, the possibilities to distribute research data directly are often very limited due to restrictions on data privacy and data confidentiality. Although these regulations are much needed, privacy constraints are also ranked among the toughest challenges to overcome in the advancement of modern day social science research [@lazer_life_2009].

Anonymizing research data might seem a quick and appealing approach to limit the unique identification of participants. However, this approach is not sufficient to fulfil contemporary privacy and confidentiality requirements [@ohm_broken_2009; @national_putting_2007]. Over the years, several other techniques have been used to increase the confidentiality of research data, such as categorizing continuous variables, top coding values above an upper bound or adding random noise to the observed values [@drechsler_synthetic_2011]. However, these methods may distort the true data relation between variables, thereby reducing the data quality and the scientific returns for re-using the same data for further research.

@@

Conceptually, the synthetic data framework is based upon the building blocks of multiple imputation of missing data, as proposed by @rubin_multiple_1987. Instead of replacing just the missing values with multiple draws from the posterior predictive distribution, one could easily *overimpute* any observed sensitive values. Similarly to multiple imputation of missing data, the multiple synthetic datasets allow for correct statistical inferences, **regardless of the fact that the analyses do not use the "true" value**. The analyses over multiple synthetic data sets should be pooled into a single inference, so that the researcher can draw valid conclusions from the pooled results. To that respect, the variance should reflect the added variability that is induced by the imputation procedure.

Potentially, this approach could fulfill the needs for openly accessibly data, without running into barriers with regard to privacy and confidentiality constraints. However, there is no such thing as a free lunch: data collectors have to put effort in creating high-quality synthetic data. Also, the quality of the synthetic data is highly dependent on the imputation models, and using flawed models to generate synthetic data might bias subsequent analyses. Conversely, if the models used to create the synthetic data are able to preserve the relationships between the variables as in the original data, the synthetic data can be nearly as informative as the observed data. Thus, to fully exploit the benefits of synthetic data, the effort to actually create these high-quality data sets should be kept at a minimum.

@@

For example, overimputing synthetic values for both the observed and missing cells in the `mice::boys` data set into the object `syn`, given the same imputation parameters as before, can be realized by the following code execution.

@@

# Simulation setup
- boys data met bootstrap om sampling variance er in te brengen
- wat evalueren we: bias | standard error | coverage
- welke pooling: T = Ubar + B/m
# psych::describe
# correlation matrix
# wgt ~ age + bmi | hgt ~ age + wgt | gen ~ age + 
# psych::describe - univariate ok
map --> psych::describe() --> Reduce("+", .) in de pipe
